---
title: "Maternal Mortality in New York"
output:  md_document:
    variant: github_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Scoping the Problem
##Maternal Mortality Rates: New York, New York City, and U.S. 

###I'm interested in examining Maternal Mortality Rates in counties across New York State, with special emphasis on the five boroughs of New York City. Here I'll create a maternal mortality rate, exmaine trends, and visualize the data with maps. 

### Loading libraries
```{r libaries}
library(tidyverse)
library(tidycensus)
library(dplyr)
library(janitor)
library(ggplot2)
library(tmap)
library(tmaptools)
library(sf)
library(leaflet)
library(plyr)
```

###Reading in Maternal Mortality Data from https://www.health.ny.gov/statistics/chac/birth/b33.htm 
```{r}
maternal_deaths <- read_csv("/Users/cecilialow-weiner/Desktop/Data Incubator Project /2013_maternal_mortality_rate.csv")
```

##Calculate Maternal Mortality
### Maternal Mortality is death within 42 dats of pregnancy termination due to complications of pregnancy, childbirth, and the peurperium. The Rate of Maternal Mortality is calculated by the number of maternal deaths in a specified geographic area divded by total resident live births for the same geographic area within a specified time, multiplied by 100,000. (Source: CDC, 2007)

```{r}
maternal_mortality_rate <- maternal_deaths %>%
  group_by(boro_name) %>%
  mutate(rate=(Maternal_deaths/Live_Births)*100000) 
```

###Scatterplot of Maternal Mortality Rate by Borough
```{r}
ggplot(data = maternal_mortality_rate) + 
  geom_point(mapping = aes(x = Live_Births, y = Maternal_deaths, color=boro_name))
```

#### Bar chart of Maternal Mortality Rate by Borough
```{r}
ggplot(data=maternal_mortality_rate, aes(x=boro_name, y=rate, fill=boro_name)) +
  geom_bar(stat="identity") +
scale_fill_brewer(palette = "Set2", name = element_blank()) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Borough Name",
    y = "Maternal Mortality Rate",
    title = "Maternal Mortality Rate by Borough")
     +
  theme_bw()
```

## Mapping Maternal Mortality in New York City

###Simple chloropeth map of maternal mortality rate in NYC

```{r}
borough <- "/Users/cecilialow-weiner/Downloads/Borough Boundaries/geo_export_838ba8d5-de6d-49a7-a7a6-681c253a5cf5.shp"
nyc_boroughs <- read_shape(file=borough, as.sf = TRUE)
qtm(nyc_boroughs)


merged_borough <- merge(nyc_boroughs,maternal_mortality_rate,by="boro_name")


tm_shape(merged_borough) +
  tm_fill(
    col = "rate",
    n = 5,
    style = "jenks"
  )

```

###Interactive map of maternal mortality rate by borough

```{r}
ttm()
tm_shape(merged_borough) +
  tm_fill(
    col = "rate",
    palette = "GnBu",
    contrast = c(0.2, 0.8),
    n = 5,
    style = "jenks",
    title = "Maternal Mortality Rate",
    legend.format = list(
      fun = function(x) {
        paste0(formatC(x, digits = 1 , format = "f"), "%")
      }
    ),
    popup.vars = c(
      "Maternal Mortality Rate" = "rate"
    ),
    id = "NAME",
    popup.format = list(rate = list(format = "f"))
  ) +
  tm_borders(col = "gray40", lwd = 1, lty = "solid", alpha = NA)
```

## Scoping the problem in New York State

```{r}
counties <- ("/Users/cecilialow-weiner/Downloads/cugir-007865/cty036.shp")
ny_counties <- read_shape(file=counties, as.sf = TRUE)
qtm(nyc_counties)


maternal_mortality_ny <- read_csv("/Users/cecilialow-weiner/Desktop/Data Incubator Project /Maternal_deaths_2013.csv")
```

###Calculate maternal mortality rate for New York Counties
```{r}
maternal_mortality_rate_ny <- maternal_mortality_ny %>%
  group_by(NAME) %>%
  mutate(rate=(Maternal_Mortality/Live_Births)*100000) 

maternal_mortality_rate_ny$COUNTYFP <- as.factor(maternal_mortality_rate_ny$COUNTYFP)

merged_counties <- merge(ny_counties,maternal_mortality_rate_ny,by="NAME")

```

## Mapping Maternal Mortality in New York State

### Simple Map
```{r}
tm_shape(merged_counties) +
  tm_fill(
    col = "rate",
    n = 5,
    style = "jenks"
  )
```

### Interactive Map

```{r}
ttm()
tm_shape(merged_counties) +
  tm_fill(
    col = "rate",
    palette = "GnBu",
    contrast = c(0.2, 0.8),
    n = 5,
    style = "jenks",
    title = "Maternal Mortality Rate",
    legend.format = list(
      fun = function(x) {
        paste0(formatC(x, digits = 1 , format = "f"), "%")
      }
    ),
    popup.vars = c(
      "Maternal Mortality Rate" = "rate"
    ),
    id = "NAME",
    popup.format = list(rate = list(format = "f"))
  ) +
  tm_borders(col = "gray40", lwd = 1, lty = "solid", alpha = NA)
```

## Scoping the Problem by Years
### United States from 1990 to 2015

### Bring in international data on maternal mortality from world bank
```{r}
international_maternal_mortality <- read_csv("/Users/cecilialow-weiner/Desktop/Data Incubator Project /international_maternal_mortality_rates.csv")

international_cleaned_names <- clean_names(international_maternal_mortality) 

international_cleaned <- c("country_name", "country_code", "x2013")

new_international_data <- international_cleaned_names[international_cleaned]

```
### Create line graph of maternal mortality rate in U.S. from 1990 to 2015

```{r}
united_states_1990_2015 <- international_cleaned_names %>%
filter(country_name=="United States")
  
us_mm_1990_2015 <- read_csv("/Users/cecilialow-weiner/Desktop/Data Incubator Project /maternal_mortality_US_1990_2015.csv") %>%
  clean_names()

ggplot(data=us_mm_1990_2015, aes(x=year, y=maternal_mortality_rate_per_100_000_live_births))+
  geom_line()+
  geom_point()
```

## New York City from 2006-2015  
```{r}
MM_Rate_NYC_time <- read_csv("/Users/cecilialow-weiner/Desktop/Data Incubator Project /maternal_mortality_NYC_2006_2015.csv")%>%
  clean_names()
```

### Create line graph of maternal mortality rate in New York City from 2006 to 2015
```{r}
ggplot(data=MM_Rate_NYC_time, aes(x=year, y=maternal_mortality_rate_per_100_000_live_births, group=1)) +
  geom_line()+
  geom_point()+
scale_x_continuous(name="Year", limits=c(2006, 2015),breaks=seq(2006,2015,1)) +
  scale_y_continuous(name="Maternal Mortality Rate", limits=c(10, 40),breaks=seq(10,40,2))
```

